Ultrafast manipulation of Heisenberg exchange and Dzyaloshinskii–Moriya interactions in antiferromagnetic insulators

### 2019-01-31 Paper/paper.tex
derive spin ham first

\begin{align}
\label{MKMH}
\hat{H}_0 &= - \sum_{\langle i,j \rangle, \sigma} t_1\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \n \\
	&-\sum_{\langle \langle i,j \rangle \rangle, \sigma}(t_2 - i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma})\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + 
	\text{U}\hat{D}
\end{align}

Where $\hat{c}_{i \sigma}^\dagger$ ($ \hat{c}_{i \sigma}$) creates (anihilates) an electron at site $i$ in spin state $\sigma$. $\hat{D} = \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ is the doublon number operator and $\text{U}$ is the on-site interaction. $t_1$, $t_2$ are the hopping amplitudes originating from both kinetic hopping. Where $\Delta$ is the intrinsic spin orbit coupling constant. $\nu_{ij}=\pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$). $\sigma^{z}$ is the third Pauli matrix. We will assume a half filling system in which the strength of the on-site interaction $\text{U}$ is much larger than the hopping amplitudes, therefore the zero double occupancies subspace $d=0$ can be taken as the low energy subspace in which the effective Hamiltonian will act.

@@

Where $t_{ij}^{\sigma}$ are the corresponding amplitudes for the unperturbed Hamiltonian, i.e. $t_{ij}^{\sigma} = t_1$ for $i,j$ being NN and $t_{ij}^{\sigma} = t_2 + i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}$ for $i,j$ being NNN. We can obtain the spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
and $n_{i \uparrow} + n_{i \downarrow} = 1$. 
and summing over the spin states. The resulting effective spin Hamiltonian is:

\begin{align}
\label{MKMHeff}
&\hat{H}_{\text{eff}} = \sum_{\langle i,j \rangle} J_{1,ij}\bs{S}_i\bs{S}_j +\n \\
&+ \sum_{\langle \langle i,j \rangle \rangle} \left\{ J_{2,ij}\bs{S}_i\bs{S}_j + \bs{D}_{2,ij} \bs{S}_i \times \bs{S}_j + \bs{S}_i \bs{\Gamma}_{ij} \bs{S}_j \right\}
\end{align}

@@

Where in the second step we added the contributions of $\langle i,j \rangle$ and $\langle j,i \rangle$ and where $J_{ij} = \frac{2t^2\text{U}}{\text{U}^2-(\epsilon_j-\epsilon_i)^2}$. This model is relevant for studying many-body localization phenomena \cite{Protopopov}.
\end{subsection}

>>>>

\begin{align}
\label{MKMH}
\hat{H}^0 &= - \sum_{\langle i,j \rangle, \sigma} t_1\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \n \\
	&-\sum_{\langle \langle i,j \rangle \rangle, \sigma}(t_2 - i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma})\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + 
	\text{U}\hat{D}
\end{align}

Where $\hat{c}_{i \sigma}^\dagger$ ($ \hat{c}_{i \sigma}$) creates (anihilates) an electron at site $i$ in spin state $\sigma$. $\hat{D} = \sum_{i=1}^M \hat{n}_{i\uparrow}\hat{n}_{i\downarrow}$ is the doublon number operator and $\text{U}$ is the on-site interaction. $t_1$, $t_2$ are the hopping amplitudes originating from both kinetic hopping. Where $\Delta$ is the intrinsic spin orbit coupling constant. $\nu_{ij}=\pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$). $\sigma^{z}$ is the third Pauli matrix. We will assume a half filling system in which the strength of the on-site interaction $\text{U}$ is much larger than the hopping amplitudes. In the strong coupling limit any state with a nonzero number of double occupancies ($d /neq 0$) will have much larger energy than those with $d=0$. We can therfore obtain an effective Hamiltonian acting on the $d=0$ subspace by standard second order perturbation techniques in the hopping terms. We can then obtain the effective spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
Following this procedure we obtain the following effective spin Hamiltonian:

\begin{align}
\label{MKMHeff0}
&\hat{H}_{\text{eff}}^0 = \sum_{\langle i,j \rangle} J_{1,ij}^0\bs{S}_i\bs{S}_j +\n \\
&+ \sum_{\langle \langle i,j \rangle \rangle} \left\{ J_{2,ij}^0\bs{S}_i\bs{S}_j + \bs{D}_{2,ij}^0 \bs{S}_i \times \bs{S}_j + \bs{S}_i \bs{\Gamma}_{ij}^0 \bs{S}_j \right\}
\end{align}

\begin{align*}
J_{1,ij}^0 &= 2t_1^2 \\
J_{2,ij}^0 &= 2t_2^2 \\
\bs{D}_{2,ij}^0 &= - 4\nu_{ij} t_2 \Delta \hat{e}_z \\
\bs{\Gamma}_{2,ij}^0 &= 2\Delta^2 \text{diag}(-1,-1,1)
\end{align*}

@@

Where $t_{ij}^{\sigma}$ are the corresponding amplitudes for the unperturbed Hamiltonian, i.e. $t_{ij}^{\sigma} = t_1$ for $i,j$ being NN and $t_{ij}^{\sigma} = t_2 + i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}$ for $i,j$ being NNN. We can obtain the spin Hamiltonian by using the relations \ref{SpinOperatorInv1} \ref{SpinOperatorInv2} and summing over the spin states. The resulting effective spin Hamiltonian is \ref{MKMHeff0} with renormalized coupling constants:

@@

Where in the second step we added the contributions of $\langle i,j \rangle$ and $\langle j,i \rangle$ and where $J_{ij} = \frac{2t^2\text{U}}{\text{U}^2-(\epsilon_j-\epsilon_i)^2}$. This model is relevant for studying many-body localization phenomena \cite{Protopopov2018}.
\end{subsection}
